---
title: "Analysis - general and PCA"
author: "Marina Paldauf"
date: "20/05/2021"
output:
  html_document:
    toc: yes
    toc_depth: 2
    float_toc: yes
    urlcolour: blue
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RMySQL)
library(reshape2)
library(GGally)
library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(cowplot)
library(corrplot)
library(devtools)
library(ggbiplot)
library(purrr)
library(viridis)
library(lubridate)
library(gridExtra)
library("FactoMineR")
library("factoextra")
library(normalr)
library(rgl)
library(rstatix)
library(readr)
library(forecast)
library(fpp2)
library(TTR)
```

## Getting Data Markova

Data access - Relevant data saved in the form of the file for privacy reasons, preventing sharing direct access to the database. 

Data set includes data form two sensors located at the Mrakova Farm near Bled, Slovenia. Sensor 1 is located in the 1st floor, and sensor 2 is located in the ground floor. Data set includes dates from 24.6.2020 until 31.12.2020. Both of them are hanging in the air attached to a string. Prediction is done on the dataset between 14.8.2020 and 31.12.2020, due to a large chunk of missing data in July 2020.


```{r data, warning=FALSE}
d.mrakova.main <- read.csv("sensor_data_2020.csv")

d.mrakova.main.predict <- d.mrakova.main %>% 
  mutate(date = as.Date(created_at, "%Y-%m-%d %H:%M:%S")) %>%
  filter(date <= as.Date("2020-12-31 23:59:59") | date <= as.Date("2020-08-14 00:00:00"))

```

Glossary of variable names:

```{r glossary, echo=FALSE, warning=FALSE}
column.names <- colnames(d.mrakova.main , do.NULL = TRUE, prefix = "col")


glossary <- cbind(column.names,c("X","sensor.calss.1", "sensor.class.2", "???", "IP", "received.signal.strength.indicator", "name", "???", "temperature", "relative.humidity", "carbon.dioxide", "pressure", "ambient.light", "volatile.organic.compounds", "particulate.matter.1", "particulate.matter.2_5", "particulate.matter.4", "particulate.matter.10", "time.created"))

glossary
```

#Cleaning data

Two datasets - one for the principal component analysis (considering only PM10 as a representative of PM variables, due to high correlated) and one for the descriptive statistics,considering all 10 variables. 

Data transformations.


```{r cleaning}

#descriptive statistics dataset

CO2 <- log(d.mrakova.main$CO2)
p <- log(d.mrakova.main$p)
ambient_light <- log(d.mrakova.main$ambient_light)
VOC_index <- log(d.mrakova.main$VOC_index)
RH <- log(d.mrakova.main$RH)
T <- log(d.mrakova.main$T)

PM1 <- log(d.mrakova.main$PM1)
PM2_5 <- log(d.mrakova.main$PM2_5)
PM4 <- log(d.mrakova.main$PM4)
PM10 <- log(d.mrakova.main$PM10)

d.mrakova.clean.all <-d.mrakova.main %>% select(X, source, meta_sn, T, RH,VOC_index, PM1, PM2_5, PM4, PM10, CO2, p, ambient_light, created_at)

#d.mrakova.main
#d.mrakova.data.all <- cbind(d.mrakova.clean.all, CO2, p, ambient_light)

d.mrakova.data.all <- d.mrakova.clean.all %>% 
  arrange(desc(p))

#d.mrakova.data.all

# remove outliers
d.mrakova.data.all <- d.mrakova.data.all[-c(1,2), ]
d.mrakova.all <- d.mrakova.data.all

d.mrakova.all$CO2[which(!is.finite(d.mrakova.all$CO2))] <- 0
d.mrakova.all$VOC_index[which(!is.finite(d.mrakova.all$VOC_index))] <- 0
d.mrakova.all$p[which(!is.finite(d.mrakova.all$p))] <- 0
d.mrakova.all$ambient_light[which(!is.finite(d.mrakova.all$ambient_light))] <- 0
d.mrakova.all$PM1[which(!is.finite(d.mrakova.all$PM1))] <- 0
d.mrakova.all$PM2_5[which(!is.finite(d.mrakova.all$PM2_5))] <- 0
d.mrakova.all$PM4[which(!is.finite(d.mrakova.all$PM4))] <- 0
d.mrakova.all$PM10[which(!is.finite(d.mrakova.all$PM10))] <- 0

#PCA dataset
d.mrakova.clean <- d.mrakova.main %>% select(X, date, source, meta_sn, created_at) %>%
  mutate(year = year(date)) %>%
  mutate(month = month(date)) %>%
  mutate(day = day(date)) %>%
  mutate(hour = as.numeric(strftime(created_at, "%H"))) %>%
  mutate(week = week(date))

d.mrakova.data <- cbind(d.mrakova.clean, CO2, p, ambient_light, VOC_index, PM10,T, RH)
d.mrakova.data <- d.mrakova.data %>% 
  arrange(desc(p))

d.mrakova.data$CO2[which(!is.finite(d.mrakova.data$CO2))] <- 0
d.mrakova.data$VOC_index[which(!is.finite(d.mrakova.data$VOC_index))] <- 0
d.mrakova.data$p[which(!is.finite(d.mrakova.data$p))] <- 0
d.mrakova.data$ambient_light[which(!is.finite(d.mrakova.data$ambient_light))] <- 0
d.mrakova.data$PM10[which(!is.finite(d.mrakova.data$PM10))] <- 0
d.mrakova.data$RH[which(!is.finite(d.mrakova.data$RH))] <- 0
d.mrakova.data$T[which(!is.finite(d.mrakova.data$T))] <- 0

# remove outliers
d.mrakova.data <- d.mrakova.data[-c(1,2), ]
d.mrakova <- d.mrakova.data

#prediction dataset

d.mrakova.main.predict <- d.mrakova.main.predict %>% select(X, date, source, meta_sn, T, RH, created_at) %>%
  mutate(year = year(date)) %>%
  mutate(month = month(date)) %>%
  mutate(day = day(date)) %>%
  mutate(hour = as.numeric(strftime(created_at, "%H"))) %>%
  mutate(week = week(date))

d.mrakova.predict <- cbind(d.mrakova.main.predict, CO2, p, ambient_light, VOC_index, PM10)

d.mrakova.predict <- d.mrakova.predict %>% 
  arrange(desc(p))

d.mrakova.predict$CO2[which(!is.finite(d.mrakova.predict$CO2))] <- 0
d.mrakova.predict$VOC_index[which(!is.finite(d.mrakova.predict$VOC_index))] <- 0
d.mrakova.predict$p[which(!is.finite(d.mrakova.predict$p))] <- 0
d.mrakova.predict$ambient_light[which(!is.finite(d.mrakova.predict$ambient_light))] <- 0
d.mrakova.predict$PM10[which(!is.finite(d.mrakova.predict$PM10))] <- 0

# remove outliers
d.mrakova.predict <- d.mrakova.predict[-c(1,2), ]

```


# Descriptive statistics

### Summary statistics of complete data set - absolute values

```{r descriptive1, echo=FALSE, warning=FALSE}

summary(d.mrakova.data.all)

```

### Summary statistics of PCA data set - logarithmic values (except Temperature and Relative Humidity)

```{r descriptive2, echo=FALSE, warning=FALSE}

summary(d.mrakova)

```

### Summary statistics of Prediction data set - logarithmic values (except Temperature and Relative Humidity)

```{r descriptive3.1, echo=FALSE, warning=FALSE}

summary(d.mrakova.predict)

```

### Missing values

```{r missing, echo=FALSE, warning=FALSE}

d.mrakova.na <- d.mrakova.data.all

#count by rows
d.mrakova.na$num_nulls <- apply(d.mrakova.na, 1, function(x) sum(is.na(x)))

#count by column
data.na <- d.mrakova.na %>%
    map_df(function(x) sum(is.na(x))) %>%
    gather(feature, num_nulls) %>%
    print(n = 100)

#table
#data.na

```

There are no missing measurements for the measured dates.


### Duplicates

Approximately 16.2% of observations are duplicated by date/time. 

```{r descriptive3, echo=FALSE, warning=FALSE}

d.mrakova.dup.all <- d.mrakova.all
d.mrakova.dup <- d.mrakova
d.mrakova.dup.predict <- d.mrakova.predict

# total number of rows - observations
count.all <- sum(complete.cases(d.mrakova.dup))

# count duplicates by date
count.d <- sum(table(d.mrakova.dup$created_at)-1)

# proportion of duplicates
count.d/count.all

# remove duplicates for further analysis
d.mrakova.non.dup.all <- d.mrakova.dup.all %>% distinct(created_at, .keep_all = TRUE)
d.mrakova.non.dup <- d.mrakova.dup %>% distinct(created_at, .keep_all = TRUE)
d.mrakova.non.dup.predict <- d.mrakova.dup.predict %>% distinct(created_at, .keep_all = TRUE)

```

```{r outliers2}

#complete dataset

outliers.no <- d.mrakova.non.dup.all

# removing outliers

#outliers.no <- d.mrakova.non.dup.all
nrow(outliers.no)

#VOC index
Q <- quantile(outliers.no$VOC_index, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$VOC_index)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(VOC_index >= low & VOC_index <= up)


#CO2
Q <- quantile(outliers.no$CO2, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$CO2)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(CO2 >= low & CO2 <= up)


#p
Q <- quantile(outliers.no$p, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$p)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(p >= low & p <= up)


#RH
Q <- quantile(outliers.no$RH, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$RH)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(RH >= low & RH <= up)


#PM1
Q <- quantile(outliers.no$PM1, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$PM1)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(PM1 >= low & PM1 <= up)


#PM2_5
Q <- quantile(outliers.no$PM2_5, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$PM2_5)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(PM2_5 >= low & PM2_5 <= up)


#PM4
Q <- quantile(outliers.no$PM4, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$PM4)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(PM4 >= low & PM4 <= up)


#PM10
Q <- quantile(outliers.no$PM10, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$PM10)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(PM10 >= low & PM10 <= up)


#ambient_light
Q <- quantile(outliers.no$ambient_light, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$ambient_light)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(ambient_light >= low & ambient_light <= up)

nrow(outliers.no)

d.mrakova.non.dup.all <- outliers.no %>% ungroup()


# PCA dataset

outliers.no <- d.mrakova.non.dup

# removing outliers

nrow(outliers.no)

#VOC index
Q <- quantile(outliers.no$VOC_index, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$VOC_index)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(VOC_index >= low & VOC_index <= up)


#CO2
Q <- quantile(outliers.no$CO2, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$CO2)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(CO2 >= low & CO2 <= up)


#p
Q <- quantile(outliers.no$p, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$p)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(p >= low & p <= up)


#RH
Q <- quantile(outliers.no$RH, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$RH)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(RH >= low & RH <= up)


#PM10
Q <- quantile(outliers.no$PM10, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$PM10)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(PM10 >= low & PM10 <= up)


#ambient_light
Q <- quantile(outliers.no$ambient_light, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$ambient_light)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(ambient_light >= low & ambient_light <= up)

nrow(outliers.no)

d.mrakova.non.dup <- outliers.no %>% ungroup()


#prediction dataset

#complete dataset

outliers.no <- d.mrakova.non.dup.predict

# removing outliers

nrow(outliers.no)

#VOC index
Q <- quantile(outliers.no$VOC_index, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$VOC_index)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(VOC_index >= low & VOC_index <= up)


#CO2
Q <- quantile(outliers.no$CO2, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$CO2)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(CO2 >= low & CO2 <= up)


#p
Q <- quantile(outliers.no$p, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$p)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(p >= low & p <= up)


#RH
Q <- quantile(outliers.no$RH, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$RH)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(RH >= low & RH <= up)


#PM10
Q <- quantile(outliers.no$PM10, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$PM10)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(PM10 >= low & PM10 <= up)


#ambient_light
Q <- quantile(outliers.no$ambient_light, probs=c(.25, .75), na.rm = FALSE)

iqr <- IQR(outliers.no$ambient_light)

up <- Q[2]+1.5*iqr # Upper Range  
low<- Q[1]-1.5*iqr # Lower Range

outliers.no <- outliers.no %>% group_by(meta_sn) %>% filter(ambient_light >= low & ambient_light <= up)

outliers.no <- outliers.no %>% group_by(meta_sn) %>% 
  filter(ambient_light > low & ambient_light < up) %>% 
  ungroup()

nrow(outliers.no)

d.mrakova.non.dup.predict <- outliers.no

```

### t-test comparison between sensors

```{r ttest}

# T
T.test <- t.test(T ~ meta_sn, data = d.mrakova.non.dup.all)

# CO2
CO2.test <- t.test(CO2 ~ meta_sn, data = d.mrakova.non.dup.all)

# p
p.test <- t.test(p ~ meta_sn, data = d.mrakova.non.dup.all)

# RH
RH.test <- t.test(RH ~ meta_sn, data = d.mrakova.non.dup.all)

# PM1
PM1.test <- t.test(PM1 ~ meta_sn, data = d.mrakova.non.dup.all)

# PM2_5
PM2_5.test <- t.test(PM2_5 ~ meta_sn, data = d.mrakova.non.dup.all)

# PM4
PM4.test <- t.test(PM4 ~ meta_sn, data = d.mrakova.non.dup.all)

# PM10
PM10.test <- t.test(PM10 ~ meta_sn, data = d.mrakova.non.dup.all)

# VOC_index
VOC_index.test <- t.test(VOC_index ~ meta_sn, data = d.mrakova.non.dup.all)


# ambient_light
ambient_light.test <- t.test(ambient_light ~ meta_sn, data = d.mrakova.non.dup.all)


T.test
CO2.test
RH.test
p.test
VOC_index.test
ambient_light.test
PM1.test
PM2_5.test
PM4.test
PM10.test

```

### Descriptive statistics by sensor

Duplicated values are approximately 10.3% for the first sensor and 4.6% for the second sensor.

```{r descriptive4, echo=FALSE, warning=FALSE}

#splitting data based on sensors

X <- split(d.mrakova.non.dup.all, d.mrakova.non.dup.all$source)

d.mrakova1 <- X[[1]]
d.mrakova2 <- X[[2]]

#examining duplicates
#MRAKOVA 1

# total number of rows - observations
#count.all.1 <- sum(complete.cases(d.mrakova1))

# count duplicates by date
#count.d.1 <- sum(table(d.mrakova1$created_at)-1)

# proportion of duplicates
#count.d.1/count.all.1

#MRAKOVA 2

# total number of rows - observations
#count.all.2 <- sum(complete.cases(d.mrakova2))

# count duplicates by date
#count.d.2 <- sum(table(d.mrakova2$created_at)-1)

# proportion of duplicates
#count.d.2/count.all.2

```

### Summary statistics by sensor

Summary statistics sensor 1 and 2: 

```{r descriptive.sensors, echo=FALSE, warning=FALSE}
d.mrakov.non.dup.summary <- d.mrakova.non.dup.all %>% split(.$meta_sn) %>% map(summary)
#sd.mrakov.non.dup.summary

```

```{r twodatasets}

d.mrakova.non.dup.1 <- d.mrakova1 %>%
  select(source,meta_sn, T, RH, CO2, p, ambient_light, VOC_index, PM1, PM2_5, PM4, PM10, created_at) %>% 
  distinct(created_at, .keep_all = TRUE) 

d.mrakova.non.dup.1$time <- as.Date(strptime(d.mrakova.non.dup.1$created_at, "%Y-%m-%d %H:%M:%S"))

#d.mrakova1

d.mrakova.non.dup.2 <- d.mrakova2 %>%
  select(source, meta_sn, T, RH, CO2, p, ambient_light, VOC_index, PM1, PM2_5, PM4, PM10, created_at) %>% 
  distinct(created_at, .keep_all = TRUE) 

d.mrakova.non.dup.2$time <- as.Date(strptime(d.mrakova.non.dup.2$created_at, "%Y-%m-%d %H:%M:%S"))

#d.mrakova2
```


### Density of individual variables for both senors

```{r descriptive.sensors2, echo=FALSE, warning=FALSE}

par(mfrow=c(2,2))

#temperature sensor 1

m<-mean(d.mrakova.non.dup.1$T);std<-sqrt(var(d.mrakova.non.dup.1$T))
hist(d.mrakova.non.dup.1$T,prob=T,main="Temperature - Sensor 1")
curve(dnorm(x, mean=m, sd=std), col="darkblue", lwd=2, add=TRUE) 

#temperature sensor 2

m2<-mean(d.mrakova.non.dup.2$T);std2<-sqrt(var(d.mrakova.non.dup.2$T))
hist(d.mrakova.non.dup.2$T,prob=T,main="Temperature - Sensor 2")
curve(dnorm(x, mean=m2, sd=std2), col="darkblue" , lwd=2, add=TRUE) 


#RH sensor 1

m3<-mean(d.mrakova.non.dup.1$RH);std3<-sqrt(var(d.mrakova.non.dup.1$RH))
hist(d.mrakova.non.dup.1$RH,prob=T,main="RH - Sensor 1")
curve(dnorm(x, mean=m3, sd=std3), col="darkblue", lwd=2, add=TRUE) 

#RH sensor 2

m4<-mean(d.mrakova.non.dup.2$RH);std4<-sqrt(var(d.mrakova.non.dup.2$RH))
hist(d.mrakova.non.dup.2$RH,prob=T,main="RH - Sensor 2")
curve(dnorm(x, mean=m4, sd=std4), col="darkblue", lwd=2, add=TRUE)

par(mfrow=c(2,2))


# CO2 sensor 1
co2 <- d.mrakova.non.dup.1$CO2

m5<-mean(d.mrakova.non.dup.1$CO2);std5<-sqrt(var(d.mrakova.non.dup.1$CO2))
hist(co2,prob=T,main="CO2 - Sensor 1", xlab="C02")
curve(dnorm(x, mean=m5, sd=std5), col="darkblue", lwd=2, add=TRUE) 

# CO2 sensor 2
co2 <- d.mrakova.non.dup.2$CO2

m6<-mean(d.mrakova.non.dup.2$CO2);std6<-sqrt(var(d.mrakova.non.dup.2$CO2))
hist(co2,prob=T,main="CO2 - Sensor 2", xlab="CO2")
curve(dnorm(x, mean=m6, sd=std6), col="darkblue", lwd=2, add=TRUE)



# ph sensor 1

m7<-mean(d.mrakova.non.dup.1$p);std7<-sqrt(var(d.mrakova.non.dup.1$p))
hist(d.mrakova.non.dup.1$p,prob=T,main="pressure - Sensor 1")
curve(dnorm(x, mean=m7, sd=std7), col="darkblue", lwd=2, add=TRUE) 

# ph sensor 2

m8<-mean(d.mrakova.non.dup.2$p);std8<-sqrt(var(d.mrakova.non.dup.2$p))
hist(d.mrakova.non.dup.2$p,prob=T,main="pressure - Sensor 2")
curve(dnorm(x, mean=m8, sd=std8), col="darkblue", lwd=2, add=TRUE) 


par(mfrow=c(2,2))

# ambient light sensor 1

m9<-mean(d.mrakova.non.dup.1$ambient_light);std9<-sqrt(var(d.mrakova.non.dup.1$ambient_light))
hist(d.mrakova.non.dup.1$ambient_light,prob=T,main="Ambient light - Sensor 1")
curve(dnorm(x, mean=m9, sd=std9), col="darkblue", lwd=2, add=TRUE) 

# ambient light sensor 2

m10<-mean(d.mrakova.non.dup.2$ambient_light);std10<-sqrt(var(d.mrakova.non.dup.2$ambient_light))
hist(d.mrakova.non.dup.2$ambient_light,prob=T,main="Ambient light - Sensor 2")
curve(dnorm(x, mean=m10, sd=std10), col="darkblue", lwd=2, add=TRUE)


# VOC index sensor 1
voc <- d.mrakova.non.dup.1$VOC_index

m11<-mean(d.mrakova.non.dup.1$VOC_index);std11<-sqrt(var(d.mrakova.non.dup.1$VOC_index))
hist(d.mrakova.non.dup.1$VOC_index,prob=T,main="VOC index - Sensor 1", xlab="VOC index")
curve(dnorm(x, mean=m11, sd=std11), col="darkblue", lwd=2, add=TRUE) 

# VOC index sensor 2

voc <- d.mrakova.non.dup.2$VOC_index

m12<-mean(d.mrakova.non.dup.2$VOC_index);std12<-sqrt(var(d.mrakova.non.dup.2$VOC_index))
hist(voc,prob=T,main="VOC index - Sensor 2", xlab="VOC index")
curve(dnorm(x, mean=m12, sd=std12), col="darkblue", lwd=2, add=TRUE)


par(mfrow=c(2,2))

# PM1 sensor 1

m13<-mean(d.mrakova.non.dup.1$PM1);std13<-sqrt(var(d.mrakova.non.dup.1$PM1))
hist(d.mrakova.non.dup.1$PM1,prob=T,main="PM 1 - Sensor 1")
curve(dnorm(x, mean=m13, sd=std13), col="darkblue", lwd=2, add=TRUE) 

# PM1 sensor 2

m14<-mean(d.mrakova.non.dup.2$PM1);std14<-sqrt(var(d.mrakova.non.dup.2$PM1))
hist(d.mrakova.non.dup.2$PM1,prob=T,main="PM 1 - Sensor 2")
curve(dnorm(x, mean=m14, sd=std14), col="darkblue", lwd=2, add=TRUE)


# PM2_5 sensor 1

m15<-mean(d.mrakova.non.dup.1$PM2_5);std15<-sqrt(var(d.mrakova.non.dup.1$PM2_5))
hist(d.mrakova.non.dup.1$PM2_5,prob=T,main="PM 2.5 - Sensor 1")
curve(dnorm(x, mean=m15, sd=std15), col="darkblue", lwd=2, add=TRUE) 

# PM2_5 sensor 2

m16<-mean(d.mrakova.non.dup.2$PM2_5);std16<-sqrt(var(d.mrakova.non.dup.2$PM2_5))
hist(d.mrakova.non.dup.2$PM2_5,prob=T,main="PM 2.5 - Sensor 2")
curve(dnorm(x, mean=m16, sd=std16), col="darkblue", lwd=2, add=TRUE) 


par(mfrow=c(2,2))

# PM4 sensor 1

m17<-mean(d.mrakova.non.dup.1$PM4);std17<-sqrt(var(d.mrakova.non.dup.1$PM4))
hist(d.mrakova.non.dup.1$PM4,prob=T,main="PM 4 - Sensor 1")
curve(dnorm(x, mean=m17, sd=std17), col="darkblue", lwd=2, add=TRUE) 

# PM4 sensor 2

m18<-mean(d.mrakova.non.dup.2$PM4);std18<-sqrt(var(d.mrakova.non.dup.2$PM4))
hist(d.mrakova.non.dup.2$PM4,prob=T,main="PM 4 - Sensor 2")
curve(dnorm(x, mean=m18, sd=std18), col="darkblue", lwd=2, add=TRUE)


# PM10 sensor 1

pm10 <- d.mrakova.non.dup.1$PM10

m19<-mean(d.mrakova.non.dup.1$PM10);std19<-sqrt(var(d.mrakova.non.dup.1$PM10))
hist(pm10,prob=T,main="PM 10 - Sensor 1", xlab="PM 10")
curve(dnorm(x, mean=m19, sd=std19), col="darkblue", lwd=2, add=TRUE) 

# PM10 sensor 2

pm10 <- d.mrakova.non.dup.2$PM10

m20<-mean(d.mrakova.non.dup.2$PM10);std20<-sqrt(var(d.mrakova.non.dup.2$PM10))
hist(pm10,prob=T,main="PM 10 - Sensor 2", xlab="PM 10")
curve(dnorm(x, mean=m20, sd=std20), col="darkblue", lwd=2, add=TRUE) 

par(mfrow=c(2,2))
```

# Corelation analysis

### Correlation matrix Mrakova 1:
(chart and table format)

```{r correlationanalysis1, echo=FALSE, warning=FALSE}

#MRAKOVA 1
d.mrakova1.data <- d.mrakova1 %>% select(T,RH,CO2, p, ambient_light, VOC_index, PM1, PM2_5, PM4, PM10)

#correlation matrix calculation and visualization
d.mrakova1.cor = cor(d.mrakova1.data)

corrplot1 <- corrplot(d.mrakova1.cor)
corrplot1

d.mrakova1.cor
```


### Correlation matrix Mrakova 2:
(chart and table format)

```{r correlationanalysis2, echo=FALSE, warning=FALSE}

#MRAKOVA 2
d.mrakova2.data <- d.mrakova2 %>% select(T,RH,CO2, p, ambient_light, VOC_index, PM1, PM2_5, PM4, PM10)

#correlation matrix calculation and visualization
d.mrakova2.cor = cor(d.mrakova2.data)

corrplot2 <- corrplot(d.mrakova2.cor)
corrplot2

d.mrakova2.cor
```
 
## PCA analysis 

#scaling and normalisation of data

```{r normandscale}

d.mrakova.non.dup.pca.norm <- d.mrakova.non.dup %>% 
  select(T, RH, VOC_index, PM10, CO2, ambient_light, p)
  
normalise(d.mrakova.non.dup.pca.norm)
scale(d.mrakova.non.dup.pca.norm)

#d.mrakova.non.dup.pca.norm

#d.mrakova.non.dup.pca.norm.all

```

```{r normandscale2, results="hide"}

#normalise(d.mrakova.non.dup.pca.norm)

#scale(d.mrakova.non.dup.pca.norm)

week <- d.mrakova.non.dup$week
meta_sn <- d.mrakova.non.dup$meta_sn

d.mrakova.non.dup.pca.norm.all <- cbind(d.mrakova.non.dup.pca.norm, meta_sn,  week)
d.mrakova.non.dup.pca.norm.all
```

PCA analysis

```{r pcadetailed}

#PCA ANALYSIS

#d.mrakova.non.dup.pca <- d.mrakova.non.dup %>% select(T, RH, VOC_index, PM10, meta_sn, CO2, ambient_light, p) %>% group_by(meta_sn)

d.mrakova.non.dup.pca <- d.mrakova.non.dup.pca.norm.all %>% group_by(meta_sn, week)

d.mrakova.non.dup.pca <- aggregate(d.mrakova.non.dup.pca[, 1:7], list(meta_sn, week), mean)

res.pca <- PCA(d.mrakova.non.dup.pca, scale.unit = TRUE, ncp = 5, graph = TRUE)

#d.mrakova.non.dup.pca <- d.mrakova.non.dup.pca.norm.all %>% 
#  group_by(d.mrakova.non.dup$meta_sn, d.mrakova.non.dup$week) %>% 
#  ggbiplot(res.pca, ellipse=TRUE, groups=d.mrakova.non.dup$meta_sn)

#graph of variables
var_plot <- ggbiplot(res.pca, group = d.mrakova.non.dup.pca$Group.1, ellipse = T) + theme_bw()
var_plot

####################################################################################################
#REMOVE THIS PART PROBABLY - DOUBLE CHECK

# EIGENVALUES
eig.val <- get_eigenvalue(res.pca)
eig.val

#graph eigne values - scree plot
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))

#RESULTS
#graph results
var <- get_pca_var(res.pca)
var


# Coordinates
head(var$coord)
# Cos2: quality on the factore map
head(var$cos2)
# Contributions to the principal components
head(var$contrib)


#QUALITY OF REPRESENTATION
head(var$cos2, 7)

#correlation plot
corrplot(var$cos2, is.corr=FALSE)

# Total cos2 of variables on Dim.1 and Dim.2
fviz_cos2(res.pca, choice = "var", axes = 1:2)


# Change the transparency by cos2 values
fviz_pca_var(res.pca, alpha.var = "cos2")


#CONTRIBUTIONS TO INDVIDIUAL VARIABLES
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)
# Contributions of variables to PC3
fviz_contrib(res.pca, choice = "var", axes = 3, top = 10)

###################################################################################
#contributions chart
fviz_pca_var(res.pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")
             )
```
## Heat Maps

Temperature

```{r heat1}
p <-ggplot(d.mrakova.non.dup, aes(day, hour, fill=T))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hourly Temps C",option ="C") +
  facet_grid(meta_sn ~ month)

p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + labs(title= paste("Hourly Temperature"), x="Day", y="Hour Commencing")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))

p
```

CO2

```{r heat2}
p <-ggplot(d.mrakova.non.dup, aes(day, hour, fill=CO2))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hourly CO2",option ="C")

p <-p + facet_grid(meta_sn ~ month)
p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + labs(title= paste("Hourly CO2"), x="Day", y="Hour Commencing")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))

p
```

VOC index

```{r heat3}
p <-ggplot(d.mrakova.non.dup, aes(day, hour, fill=VOC_index))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hourly VOC index",option ="C")

p <-p + facet_grid(meta_sn ~ month)
p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + labs(title= paste("Hourly VOC index"), x="Day", y="Hour Commencing")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))

p
```

Particulate matter (size 10)

```{r heat4}
p <-ggplot(d.mrakova.non.dup, aes(day, hour, fill=PM10))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hourly particulate matter particles 10",option ="C")

p <-p + facet_grid(meta_sn ~ month)
p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + labs(title= paste("Hourly PM10"), x="Day", y="Hour Commencing")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))

p
```

Relative humidity

```{r heat5}
p <-ggplot(d.mrakova.non.dup, aes(day, hour, fill=RH))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hourly RH",option ="C")

p <-p + facet_grid(meta_sn ~ month)
p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + labs(title= paste("Hourly RH"), x="Day", y="Hour Commencing")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))

p
```

Ambient light

```{r heat6}
p <-ggplot(d.mrakova.non.dup, aes(day, hour, fill=ambient_light))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hourly ambient light",option ="C")

p <-p + facet_grid(meta_sn ~ month)
p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + labs(title= paste("Hourly ambeint light"), x="Day", y="Hour Commencing")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))

p
```

Pressure

```{r heat7}
p <-ggplot(d.mrakova.non.dup, aes(day, hour, fill=p))+
  geom_tile(color= "white",size=0.1) + 
  scale_fill_viridis(name="Hourly pressure",option ="C")

p <-p + facet_grid(meta_sn ~ month)
p <-p + scale_x_continuous(breaks =c(1,10,20,31))
p <-p + labs(title= paste("Hourly pressure"), x="Day", y="Hour Commencing")
p <-p + theme(legend.position = "bottom")+
  theme(plot.title=element_text(size = 14))+
  theme(axis.text.y=element_text(size=6)) +
  theme(strip.background = element_rect(colour="white"))+
  theme(plot.title=element_text(hjust=0))+
  theme(axis.ticks=element_blank())+
  theme(axis.text=element_text(size=7))+
  theme(legend.title=element_text(size=8))+
  theme(legend.text=element_text(size=6))

p
```
